---
title: "Tutorial"
author: "Giada Leone"
date: "14/11/2021"
output:
  html_document: default
  pdf_document: default
---

# Creating **static and interactive** maps in R for intermediate users

This tutorial is aimed at people who have some experience in using R and would like to take their **spatial data visualisation** to the next level by using R as a GIS like software using specific packages.

R is very powerful due to its wide applications in statistics, modelling, data visualisation and spatial data. In this tutorial we will be focussing on spatial data and different ways to visualise it.

### Tutorial Aims:
1. Data preparation and wrangling for spatial visualisation
2. Creating basic maps using ggplot2
3. Creating interactive maps using the mapview package
4. ...

These are the libraries we need to load. 
If you have any troubles installing the ```sf``` package you can check out this link for suggestions: https://r-spatial.github.io/sf/ 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)  # Data wrangling
library(janitor)    # Cleaning datasets
library(ggthemes)   # Themes for ggplot2
library(sf)         # 
library(mapview)    # Interactive maps
```

Importing the dataset
```{r}
beaches <- read.csv("../data/beachwatch_data.csv", header = TRUE)
str(beaches)  # Checking data structure

```

Preparing the data
```{r}
# cleaning data frame column names
beaches <- beaches %>% clean_names() 

# deleting unused columns
beaches <- select(beaches,-seq(12, 109), -seq(112, 170))  

# dropping columns which have NA values
beaches <- drop_na(beaches)  

```

Changing data type

```{r}
beaches <- beaches %>% mutate_if(is.character, as.factor) %>% 
                       mutate_if(is.integer, as.numeric)

```


Let's use ggplot to create a map to show us all the beaches around the UK which have data from Great British Beach Clean surveys. 

Creating a basic map in ggplot2







Here we specify the coordinates (longitude and latitude) of our points within the ```aes()``` function and we also tell ggplot to colour the points depending on beach region.





```{r}
(map <- ggplot(beaches,
               aes(x = beach_longitude,
                   y = beach_latitude, 
                   colour = beach_region)) +
   borders("world", colour = "grey5", fill = "black", size = 0.2) +
   coord_cartesian(xlim = c(-10, 3), ylim = c(50.3, 59)) +
   theme_map() +
   geom_point(size = 1) +
   theme(plot.title = element_text(size = 12, hjust = 0.5),
         legend.position = c(0.77, 0.4),
         legend.box.background = element_rect(color = 'grey', size = 0.5)) +
   labs(title = "Beach Locations by Region",
        colour = "UK Regions"))
```










Spatial analysis 
```{r}
beaches_obj <- st_as_sf(beaches, coords =  c('beach_longitude', 'beach_latitude'), crs = 4326)

# examine the sf objec
beaches_obj
str(beaches_obj)

plot(beaches_obj)

plot(st_geometry(beaches_obj))
```

Interactive map 
```{r}
mapview(beaches_obj, zcol = 'beach_region')
```


